---
title: 'Unit 2 Case Study: Topology over Time'
subtitle: "ECI 589 Social Network Analysis and Education"
author: "Dr. Shaun Kellogg"
date: "`r format(Sys.Date(),'%B %e, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options:
  markdown:
    wrap: 72
bibliography: lit/references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare

Our Unit 2 Case Study: Hashtag Common Core revisits Alan Daly, Miguel
del Fresno and Christian Kolouch who examined the intense debate
surrounding the Common Core State Standards education reform as it
played out on Twitter. As noted on their expansive and interactive
website for the [#COMMONCORE
Project](https://www.hashtagcommoncore.com), the Common Core was a major
education policy initiatives of the early 21st century that sought to
strengthen education systems across the United States through a set of
specific and challenging education standards. Although these standards
once enjoyed bipartisan support, we'll see later in this walkthrough how
these standards have become a political punching bag.

In addition to investigating tweets around the these controversial state
standards, our Unit 2 Walkthrough will cover the following topics
pertaining to each data-intensive workflow process:

1.  **Prepare**: Prior to analysis, we'll take a look at the context
    from which our data came, formulate some research questions, and get
    introduced the {rtweet} R package for using the Twitter API.

2.  **Wrangle**: Wrangling data entails the work of manipulating,
    cleaning, transforming, and merging data. In section 2 we will learn
    about the {tidygraph} package for creating network objects.

3.  **Explore**: In section 3, we use the {tidygraph} package and the
    companion {ggraph} package to calculate a range of centrality
    measures and learn how to illustrate some of these stats through
    network visualization.

4.  **Model**: We wrap up our analysis in Section 4 by introducing
    community detection and sentiment analysis algorithms for
    identifying groups and gauging sentiment about the common core.

5.  **Communicate**: We briefly reflect on our walkthrough in
    preparation for our independent analysis next week.

### Review the Research

In [Social Network Analysis and Education: Theory, Methods &
Applications](https://methods.sagepub.com/book/social-network-analysis-and-education)
@carolan2014 makes the following distinctions between mathematical and
statistical approaches to social network analysis:

> **Mathematical approaches** focus on what a network of actors "looks
> like" by describing the network using sociograms and/or network
> measures such as reciprocity, centrality, and density. However, these
> approaches "tend to regard the measured relationships and their
> strengths as accurately reflecting the real, final, or equilibrium
> status of the network."

For Unit 3, we move beyond previous "mathematical approaches"

Specifically, we will try to replicate the P\* models used in Chapter 9
of @carolan2014 in order to test hypotheses about individual and network
characteristics that may explain who school leaders select for
collaboration and confidential exchanges.

#### Research Questions

For this walkthrough, we will be using school leadership data were
collected at two school districts over 3 consecutive years that was used
to answer the following research questions:

1.  Is there a relationship between the frequency of collaboration
    between school leaders and how often they turn to each other to
    discuss issues of a confidential nature?
2.  Do school leaders prefer to collaborate with those with whom they
    have collaborated in the past, or is there some other reason?
3.  Does gender or some other individual attribute predicts confidential
    exchanges between school leaders, or does some previous relation
    have a stronger effect?
4.  Does collaboration between leaders explain one's level of trust in
    one's administrative colleagues?
5.  Can we distinguish among different groups of school leaders based on
    how frequently they collaborate, and if so, are these groupings
    related to the level at which they work (school versus district)?

#### Data Collection

For each consecutive year, school district leaders were invited to
complete a survey that collected individual:

-   **Demographic information** (e.g., gender, ethnicity, marital
    status, age, years of experiences);

-   **Network relationships types** (e.g., collaboration, confidential
    exchanges, energy, expertise, leaders approached for support,
    work-related issues, input, recognition, best practices, and
    innovation);

-   **Frequency of interactions** they have with those nominated
    individuals on a four-point frequency scale ranging from 1 (the
    least frequent) to 4 (1--2 times a week);

-   **Leadership efficacy** items were designed based on the Principal
    Efficacy Scale used in Daly et al. (2011) and Tschannen-Moran and
    Gareis's (2004) studies. The efficacy scale includes 18 items rated
    on a 9-point Likert scale ranging from 1 (None at all) to 9 (A great
    deal);

-   **Trust** scale contains eight items rated on a 7-point Likert scale
    ranging from 1 (Strongly disagree) to 7 (Strongly agree) modified
    from Tschannen-Moran and Hoy (2003).

#### Analyses

Two modeling techniques unique to network analysis and demonstrated in
Chapter 9 of @carolan2014 are the quadratic assignment procedure (QAP)
and Exponential Random Graph Models (ERGM):

-   **QAP/MR**: The quadratic assignment procedure (QAP) developed by
    Hubert (1987) and Krackhardt (1987b) tests the null hypothesis of no
    correlation between the two networks and adjusts for this dependence
    between networks by repeatedly permuting the order of rows and
    columns of one of the networks while keeping the other network
    intact. The QAP is based on regression models and permutation tests
    for valued (i.e., continuous) relational variables.

-   **P1 and P\* (P-Star)**: Both of these models are some of the first
    to make use of the ERGM, which provides a basis for comparing
    whether a network's observed structural properties occur more
    frequently than you could expect from chance alone. ERGMs can be
    used to model ties in complete networks but do so in a manner that
    explains the presence (or absence) of ties as a function of
    individual-and/or network-level characteristics. While QAP and
    MR-QAP procedures control for network structure through
    permutations, these ERGMs attempt to explain it.

#### Key Findings

In response to research questions 1 & 2 from above, Carolyn reported
that "while collaboration in year 1 does not significantly predict
collaboration in year 3, confidential exchanges in year 1 does" and
suggested that "collaboration among school leaders provides an important
foundation for more sensitive, perhaps even deeper, relations (e.g.,
confidential exchanges) at a later point in time."

### 

For our Unit 3 Walkthrough, we focus specifically on generative
processes, or micro-level mechanisms, the might help explain who school
leaders select to collaborate with or engage in confidential exchanges.
Review Chapter 9 of SNA and Education and report out on the findings
from the ERGM analysis in response to Research Question 3 below:

3.  **Does gender or some other individual attribute predicts
    confidential exchanges between school leaders, or does some previous
    relation have a stronger effect?**

    -   YOUR RESPONSE HERE

Based on what you know about networks and the context so far, what other
specific question(s) might the data and analyses described allow you to
address. For example, the concept of homophily, or the tendency for
people to seek out or form ties to those who are similar to themselves,
was introduced in our readings and hypotheses about homophily can easily
be tested using using ERGMs:

-   YOUR RESPONSE HERE

Save this question for later as you'll have a chance to empirically test
it by the end of this walkthrough.

## 

Load Packages

```{r cars}
library(readxl)
library(tidyverse)
library(igraph)
library(tidygraph)
library(ggraph)
library(janitor)
library(knitr)
library(DT)
```

\

## Wrangle

You can also embed plots, for example:

### Import Data

One of our primary goals is to replicate the ERGM analysis from Chapter
9: Network Data and Statistical Models [@carolan2014]. Specifically,
we'll aim to examine an

To do so, we'll need to import two Excel files from the [Social Network
Analysis and Education companion
site](https://studysites.sagepub.com/carolan/study/resources.htm). The
first file contains our edges stored as a square matrix (more on this
later) and the second file is a standard rectangular data frame that
contains attributes for each node, i.e. school leaders. These files are
included in the data folder of your R Studio project. A description of
each file from the companion website is linked above and each data file
is linked below:

1.  [**School Leaders Data Chapter
    9_a.**](https://studysites-sagepub-com.prox.lib.ncsu.edu/carolan/study/materials/datasets/99472_ds7.xls) This
    adjacency matrix reports on "collaboration" ties among 43 school
    leaders in year 1 of a three-year study. This is a directed valued
    (weighted) network measured on five-point scale ranging from 0 to 4,
    with higher values indicating more frequent collaborations (1--2
    times/week). These data are used throughout Chapter 9.

2.  [**School Leaders Data Chapter
    9_b.**](https://studysites-sagepub-com.prox.lib.ncsu.edu/carolan/study/materials/datasets/99472_ds8.xlsx) This
    adjacency matrix reports on "collaboration" ties among 43 school
    leaders in year 3 of a three-year study. This is a directed valued
    (weighted) network measured on five-point scale ranging from 0 to 4,
    with higher values indicating more frequent collaborations (1--2
    times/week). 

Recall from above that our relations, or edges, are stored as a valued
[adjacency matrix](https://en.wikipedia.org/wiki/Adjacency_matrix) in
which columns and rows consist of the same actors and each cell contains
information about the tie between each pair of actors. In our case, the
tie is a directed and valued "arc" where the value indicates the
strength of the relationship.

Use the `read_excel()` function to import the
`School Leaders Data Chapter 9_d.xlsx` file, add an argument setting the
column names to `FALSE` since our file is a simple matrix with no header
or column names, and assign the matrix to a variable named
`leader_matrix`:

**Hint:** Type `?read_excel` into the console and check the arguments
section to find the name of the argument used to set column headers.

```{r import year }
year_1_collaboration <- read_excel("data/year-3-collaboration.xlsx", 
                            col_names = FALSE)
```

Use the code chunk below to inspect the matrix you just imported and
answer the questions that follow:

**\
**

**Hint:** Type `?read_excel` into the console and check the arguments
section to find the name of the argument used to set column headers.

```{r}
#YOUR CODE HERE
leader_matrix <- read_excel("data/School Leaders Data Chapter 9_d.xlsx", 
                            col_names = FALSE)
```

Use the code chunk below to inspect the matrix you just imported and
answer the questions that follow:

```{r}
class(year_1_collaboration)

year_1_collaboration
```

1.  Is our network a square matrix? What does this tell you about the
    number of nodes in our network?

    -   

2.  Does our network contain self-loops? How do you know?

    -   

3.  What do the values in each cell of our matrix indicate?

    -   

With that in mind, and before we can dichotomize our "matrix," we first
need to convert it to a matrix object recognized by R using the
`as.matrix()` function. We can then check to data format of our
`leader_matrix` using the `class()` function:

```{r}
year_1_matrix <- year_1_collaboration |>
  as.matrix()

class(year_1_matrix)

year_1_matrix
```

\

## Dichotomize Matrix

Since one of our goals is a crude replication of the analysis
demonstrated in Chapter 9, and since working with "valued networks" is a
little more complex than binary networks, which indicate just the
presence or absence of a relationship between two actors, we will need
to "dichotomize" our matrix. In practice this simply entails converting
our valued matrix to a binary matrix with cells containing just 1s or
0s.

In their article, [Valued Ties Tell Fewer Lies: Why Not To Dichotomize
Network Edges With Thresholds](https://arxiv.org/abs/1101.0788), Thomas
and Blitzstein (2018) cautions against dichotomization but highlight
several reasons why the procedure is appealing to social network
analysts:

1.  **Use of Exclusively Binary Methods.** Several classes of models
    have been designed to incorporate binary information directly,
    including the exponential random graph model.

2.  **Ease of Input and Data Collection.** The need to classify
    continuously-valued quantities into a set of discrete groups is
    widespread throughout all of science and technology, particularly
    because of the associated need to make clear decisions based on this
    information.

3.  **Ease of Output in Graphical Representations.** The visual appeal
    of graphs and networks has contributed to much of the field's
    attention in the past decade. When plotting a graphical structure, a
    clever choice of threshold can illuminate which nodes are most
    central, which connections the most vital.

4.  **Sparsity of Structure.** In data where there are very few natural
    zeroes (if any), di- chotomization provides a way to select for a
    small number of connections which are thought to be of the greatest
    importance to the system, or to nominate a number of ties for more
    in-depth study.

5.  **Binning To Address Nonlinearity and Reduce Noise.** If there is a
    nonlinear relationship in the data, binning the data into distinct
    ordinal categories has many advantages, namely, the reduction of
    total mean-squared error, and a corresponding increase in power for
    detecting a true non-zero relationship over an improperly specified
    linear analysis.

As the title of their article suggests, however, @thomas2011valued argue
that the motivations for dichotomization should be revisited as
dichotomization produces a range of problematic issues.

Both the collaboration and confidential help network data were
dichotomized by recoding values originally coded as a 3 and 4 recoded to
1, indicating the presence of a directed tie for both relations, and
zero otherwise.

To dichotmize our our matrix, the following code will "assign" 0's to
all values in our matrix that are less than or equal to 2, and 1's to
all values that are greater the or equal to 3:

```{r}
leader_matrix[leader_matrix <= 2] <- 0

leader_matrix[leader_matrix >= 3] <- 1
```

### Add Row and Column Names

Before we can convert to an edge-list, we will also need to add the
names of our nodes to the columns and rows of our matrix. These are
stored in the `ID` column of our `leader_node` data frame. We can use
the `$` operator to select these names and assign to our `leader_matrix`
using the `rownames()` and `colnames()` functions respectively:

```{r}
rownames(year_1_matrix) <- 1:43

colnames(year_1_matrix) <- 1:43

year_1_matrix
```

## 

### Network Object

```{r}


year_1_network <- as_tbl_graph(year_1_matrix, directed = TRUE)

year_1_network
```

### 👉 Your Turn ⤵

## Explore

### Describe the Network Visually

```{r}

ggraph(year_1_network, layout = "fr") +
  geom_node_point() +
  geom_edge_link()

```

### Describe the Network Mathematically

#### Network Size

```{r}

year_1_network

year_1_network |>
  activate(edges) |>
  mutate(size = graph_size())

year_1_network |>
  gsize()

gsize(year_1_network)

```

#### Density

\

```{r}

year_1_network |>
  edge_density()

edge_density(simplify(year_1_network))


```

####  Reciprocity

```{r}

year_1_network |>
  reciprocity()

year_1_network |>
  activate(edges) |>
  mutate(reciprocated = edge_is_mutual()) |>
  as_tibble() |>
  datatable()

```

#### Transitivity

```{r}

year_1_network |>
  transitivity()


```
