---
title: 'Unit 5 Case Study: Exposure and Influence'
subtitle: "ECI 589 Social Network Analysis and Education"
author: "Data Science in Education Using R"
date: "`r format(Sys.Date(),'%B %e, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
    code_folding: show
editor_options:
  markdown:
    wrap: 72
bibliography: lit/references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## 1. PREPARE

Our brief case study for Unit 5 is adapted from an exercise on network
influence included in the appendix of [Data Science in Education Using
R](https://datascienceineducation.com/c20.html#c20c). In this case study
we explore the process of network influence, which is the flip side of
the network selection processes we examined in our previous case study.
Both of these processes are commonly the focus of statistical analyses
of networks. Selection and influence, however, do not interact
independently and often affect each other reciprocally (Xu, Frank, &
Penuel, 2018).

Let's first quickly define these two processes:

-   **Selection**: the process of choosing relationships

-   **Influence**: the process of how our social relationships affect
    behavior

### 1a. Research Question

In Unit 5, we will briefly examine a hypothetical example of peer
influence on attitudes towards STEM careers among a small group of
friends. Specifically, we will try to answer the following question:

> Are student' attitudes towards careers in STEM influenced by the
> attitudes of their friends?

### 1b. Load Libraries

While selection and influence processes are complex, it is possible to
study them using data about people's relationships and behavior.
Happily, the use of these methods has expanded along with R. In fact,
long-standing R packages have become some of the best tools for studying
social networks. Additionally, while there are many nuances to studying
selection and influence, these are models that can be carried out with
relatively simple modeling techniques like linear regression, as we'll
see later in this case study.

#### 👉 Your Turn ⤵

To begin, load the the following packages that will be using for our
analysis of network influence:

-   tidyverse
-   tidygraph
-   ggraph
-   [moderndive](https://github.com/moderndive/moderndive#moderndive-r-package-)

```{r load-libraries}
# YOUR CODE HERE

```

------------------------------------------------------------------------

## 2. WRANGLE

To examine a hypothetical case of network selection, we'll create three
different data frames:

-   One **edgelist** that contains the *nominator* and *nominee* for a
    relationship. For example, if Stefanie says that José is her friend,
    then Stefanie is the nominator and José the nominee. Our edgelist
    will also contain variable indicating the weight, or strength, of
    their relation.

-   Two **nodelists** indicating the values of some behavior - an
    outcome - at two different time points. In our case, these values
    will represent a measure of students' attitudes towards careers in
    STEM.

### 2a. Create an Edgelist

First let's create an edgelist with three variables for a nominator,
nominee, and strength of the relation. Run the following code to create
[tibble](https://tibble.tidyverse.org) -- a tidy table or data frame --
that contains information about each friendship dyad in our network with
a corresponding strength of "1", which simply indicates the presence or
absence of a tie:

```{r, create-edgelist}

student_edges <- tibble(nominator = c(2, 1, 3, 1, 2, 6, 3, 5, 6, 4, 3, 4), 
                        nominee = c(1, 2, 2, 3, 3, 3, 4, 4, 4, 5, 6, 6), 
                        relate = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1) ) 

student_edges
```

### 2a. Create Nodelists

Next, let's create a data frame, or tibble, that contains the nominee
and the values of a behavior at the first time point. For our
hypothetical example, our `time_1` variable represents a composite score
for each students' attitude towards STEM careers at the beginning of the
semester and measured on a scale of -3 (very negative) to 3 (very
positive).

```{r, nominee-attitudes}
nominee_data <- tibble(nominee = c(1, 2, 3, 4, 5, 6), 
                       time_1 = c(2.4, 2.6, 1.1, -0.5, -3, -1)
                       )

nominee_data
```

We will use this data later in our Explore section to create a new
measure called an "exposure term" that captures the extent to which an
individual is "exposed" to some attribute or characteristic there alters
in the network.

#### 👉 Your Turn ⤵

Create a third data frame for data collected at the end of the semester
on students' attitudes towards STEM careers called `nominator_data`.
Your output should look something like this:

| nominator | time_2 |
|-----------|--------|
| 1         | 2.0    |
| 2         | 2.0    |
| 3         | 1.0    |
| 4         | -0.5   |
| 5         | -2.0   |
| 6         | -0.5   |

```{r, nominator-attitudes}
# YOUR CODE HERE
nominator_data <- 

nominator_data
```

### 2c. Create Network with Attributes

Recall from previous units that we introduced the {tidygraph} package
for preparing and summarizing network data. Similar to previous units,
we will use the `tbl_graph()` function to convert our `student_edges`
and `nominee_data` into a network graph object, by including the
following arguments and supplying the appropriate code:

-   `edges =` expects a data frame, in our case `student_edges`,
    containing information about the edges in the graph. The nodes of
    each edge must either be in a `to` and `from` column, or in the two
    first columns like the data frame we provided.

-   `nodes =` expects a data frame, in our case `nominee_data`,
    containing information about the nodes in the graph.

-   `directed =` specifies whether the constructed graph be directed.

```{r create-network}
# YOUR CODE HERE
student_network <- tbl_graph(edges = student_edges,
                             nodes = nominee_data,
                             directed = TRUE)


student_network
```

Congrats! You made it to the end of data wrangling section and are ready
to start analysis!

------------------------------------------------------------------------

## 3. EXPLORE

In Section 3, we use the {tidygraph} package for retrieving network
descriptives and the {ggraph} package to create a network visualization
to help illustrate these metrics. Specifically, in this section we will:

a.  **Examine Basic Descriptives**. We quick revisit calculating basic
    centrality measures for our nodes and focus primarily on creating an
    "exposure term" and summarizing each students exposure to his
    friends attitudes towards STEM careers.

b.  **Make a Sociogram**. Finally, we wrap up the explore phases by
    creating a simple network and tweak key elements like the size,
    shape, and position of nodes and edges to better at communicating
    key findings.

### 3a. Examine Descriptives

As noted in SNA and Education [@carolan2014], many analyses of social
networks are primarily descriptive and aim to either represent the
network's underlying social structure through data-reduction techniques
or to characterize network properties through network measures.

#### Calculate Node Degree

Recall that **degree** is the number of ties to and from an ego, or in
the case of a "simple graph" like ours, degree is simply the number of
people to whom someone is connected. In a directed network,
**in-degree** is the number of connections or ties received, whereas
**out-degree** is the number of connections or ties sent.

The `activate()` function from the {tidygraph} package allows us to
treat the nodes in our network object as if they were a standard data
frame to which we can then apply tidyverse functions such as `select()`,
`filter()`, and `mutate()`.

We can use the `mutate()` functions to create new variables for nodes
such as measures of degree, in-degree, and out-degree using the
`centrality_degree()` function in the {tidygraph} package.

#### 👉 Your Turn ⤵

Use the following code chunk to calculate some basic degree measures for
each student:

```{r add-degree}
#YOUR CODE HERE
student_network <- student_network |>
  

student_network
```

### 4a. Determine "Exposure"

Next we'll create an simple exposure term. The idea is behind this
concept is relatively straightforward.

> The **exposure term** captures how your interactions or relationships
> with someone over a period of time impacts an outcome.

Our simple linear regression model in the next section will test whether
changes attitudes towards STEM careers were significatnly impacted by
the attitudes of students they identified as friends.

First, let's create a new data frame called `student_data` which
combines our edgelist along with the attitudes of the `nominee`.

```{r join-time-1}

student_data <- left_join(student_edges,
                          nominee_data,
                          by = "nominee")

student_data
```

In the data frame above we see that student 2 was "exposed" to a friend
(student 1) with a positive attitude towards STEM careers measured at
2.4.

#### Calculate Exposure Term

To calculate an exposure term, we need to create a new variable, which
we'll call `exposure`, that captures the both their friend's attitude
and the strength of the relationship.

```{r calculate-exposure}
# Calculating exposure 
student_exposure <- student_data |>
  mutate(exposure = relate * time_1) 
# Calculating mean exposure 

student_exposure
```

Since the strength of each friendship was based on simply the presence
or absence of a named friend as represented by the value of "1", the
exposure term simple equals their nominee's attitude at time point 1.
However, if we had captured valued ties and asked students to indicate
the strength of each friendship, such as "best friend", our exposure
terms might be larger for some friends since those friendships may carry
more "weight" and have a great impact on the nominators' attitudes.

#### Calculate Mean Exposure

Now that we have an exposure term for each students' relationship, for
the purposes of modeling in our next section, we need to create a single
measure that attempts to quantify their overall exposure in the network.
To do so, we'll need to calculate the average `exposure` for each friend
nominated, the output of which should look like this:

| nominator | exposure_mean |
|-----------|---------------|
| 1         | 1.8500000     |
| 2         | 1.7500000     |
| 3         | 0.3666667     |
| 4         | -2.0000000    |
| 5         | -0.5000000    |
| 6         | 0.3000000     |

Since student 1 named two friends whose attitudes towards STEM careers
prior to the semester were 2.1 and 1.1 respectively, his mean "exposure"
would be (2.6+1.1)/2 = 1.85, somewhat less that their own attitude at
the start, which for student one was 2.4. We might expect, there for
that student 1's attitude might decrease by the end of the semester
since, on average, he's exposed to friends with more negative views.

#### 👉 Your Turn ⤵

Complete the following code chunk to create a `mean_exposure` data frame
that summarizes the data from our `student_exposure` data frame.

```{r}
#YOUR CODE HERE
mean_exposure <- student_exposure |>
  

mean_exposure
```

**Hint:** consider using the `group_by()` and `summarize()` functions to
help calculate the mean exposure for each student.

### 3b. Visualize Network Properties

#### 👉 Your Turn ⤵

Try creating a **sociogram** for our small student network by modifying
the code below and/or adding new ones for
[layouts](https://ggraph.data-imaginist.com/articles/Layouts.html),
[nodes](https://ggraph.data-imaginist.com/articles/Nodes.html), and
[edges](https://ggraph.data-imaginist.com/articles/Edges.html) that
highlights the following for students in our network:

-   number of name friends
-   attitudes towards STEM careers
-   corresponding label for each student

Then answer the questions that follow below.

```{r create-sociogram}
#YOUR CODE HERE

```

Based on the the mean exposure calculated above, and the sociogram you
just created, what predictions do you have about how students attitudes
might change at time 2?

-   YOUR RESPONSE HERE

Congrats! You made it to the end of the Explore section and are ready to
learn a little about modeling network influence processes using simple
linear regression!

------------------------------------------------------------------------

## 4. MODEL

Since our goal is to determine if there is a relationship between a
students' attitude towards careers in STEM based on their exposure to
their friends attitudes, we will need to a final data frame for our
model that contains their attitudes at `time_2` (our outcome and
dependent variable), their `mean_exposure` (our predictor variable),
while also controlling for their attitudes at `time_1`.

First, let's join our `mean_exposure` data frame with our
`nominator_data` frame using the `nominator` variable to match our value
in both data frames. Let's also rename our `nominator` variable to
`student` to avoid confusion.

```{r, join-time2}

final_data <- 
  # data3 already has nominator, so no need to change 
  left_join(mean_exposure, nominator_data, by = "nominator") |>
  rename(student = nominator)

final_data
```

#### 👉 Your Turn ⤵

Now that we have our dependent and predictor variables, let also join
our `nominee_data` that contains our `time_1` variable for each student.

```{r, join-time1}
#YOUR CODE HERE
final_data <- nominee_data |>
  

final_data
```

**Hint:** Your `final_data` frame should look something like this since
our model will examine how well students' attitudes ar `time_1` as well
as their `mean_exposure` to their named friends' attitudes predict their
attitudes at `time_2`:

| student | time_1 | mean_exposure | time_2 |
|---------|--------|---------------|--------|
| 1       | 2.4    | 1.8500000     | 2.0    |
| 2       | 2.6    | 1.7500000     | 2.0    |
| 3       | 1.1    | 0.3666667     | 1.0    |
| 4       | -0.5   | -2.0000000    | -0.5   |
| 5       | -3.0   | -0.5000000    | -2.0   |
| 6       | -1.0   | 0.3000000     | -0.5   |

#### Regression (Linear Model)

Calculating the exposure term is the most distinctive and important step
in carrying out influence models. Now, we can use a linear model to find
out how much relations - as captured by the exposure term - affect their
attitudes at the end of the semester.

Run the following code to create a simple linear regression model that
contains `time_2` as our dependent variable and `time_1` and
`mean_exposure` as our "independent" variables.

```{r}

model1 <- lm(time_2 ~ time_1 + mean_exposure, data = final_data) 

summary(model1)

get_regression_table(model1)
```

#### 👉 Your Turn ⤵

Take a look at [Chapter 6.2.2 Regression
plane](https://moderndive.netlify.app/6-multiple-regression.html#model3table)
from *Statistical Inference via Data Science: A* *ModernDive into R and
the Tidyverse* to help you provide a basic interpretation of our
findings below:

-   YOUR REPONSE HERE

------------------------------------------------------------------------

## 5. COMMUNICATE

For your final Your Turn, your goal is to distill our analysis from
above into a simple "data product" designed to illustrate key findings
about changes in the collaboration network over time:

1.  **Select.** Select a finding from our analysis above, or a new
    analysis if so motivated, that you think would be interesting or
    relevant and that helps answer our research question.

2.  **Polish.** Create and polish a data visualization and/or data table
    to communicate your selected findings.

3.  **Narrate.** Write a brief narrative (2-3 paragraphs) to accompany
    your visualization and/or table that includes the following:

    -   The question or questions guiding the analysis;

    -   The conclusions you've reached based on our findings;

    -   How your audience might use this information;

    -   How you might revisit or improve upon this analysis in the
        future.

#### 👉 Your Turn ⤵

Use the code chunk below create a polished table and/or visualization(s)
and write a brief narrative in the space that follows.

### Data Visualization or Table

```{r create_data_product}
# YOUR CODE HERE


```

### Narrative

NARRATIVE GOES HERE...

### 🧶 Knit & Check ✅

Congratulations - you've completed the Unit 5 case study! To share your
work, click the drop down arrow next to the ball of yarn that says
"Knit" at the top of this markdown file, then select "Knit top HTML".
Assuming your code contains no errors, this will create a web page in
your Files pane that serves as a record of your work.

Once your file has been knitted, you can publish this file online using
RPubs, or share the HTML file through another means.

### References
